\chapter*{Acknowledgments} \label{chapter:acknowledgements}
\addcontentsline{toc}{chapter}{\nameref{chapter:acknowledgements}}

First and foremost, I would like to express my sincere gratitude to my supervisors Prof. Dr. Dirk Lamoen and Prof. Dr. Bart Partoens for providing continuous support, guidance and insightful feedback and discussion during the past four years. Second, I want to thank the many colleagues and collaborators I have had the pleasure to work with during my research, their valuable perspectives have been of great benefit to my work. A special mention here should go to Dr. Brandon Wood, for giving me the opportunity to work at Lawrence Livermore National Laboratory during the computational chemistry and materials science summer institute. My appreciation also extends to the academic and technical staff of the University of Antwerp and the Flemish Supercomputer service. Without their infrastructure and hard work, my research would not have been possible.

On a more personal note, I would like to thank my friends and family for their encouragement and contributions. Specifically, my thanks goes to Nils Deputter, who designed the cover of my thesis as well as some of the images from the chapter on solar cells. I am also very grateful to Elise Stynen, for always supporting me during my PhD and putting up with the long hours at work and times when stress or fatigue made my demeanor less than pleasant. Thank you for your understanding and love. Last but not least, I would like to thank my mother Tilda Stockmans. Her incredible strength and perseverance in face of the many hardships she has had to endure has always inspired me. It would not be an exaggeration to say that a large part of my motivation to succeed comes from a desire to make her proud. To make sure she knows that, no matter what has happened in the past, she has always been a wonderful mother, and the person I respect most in this world.
