\chapter*{Conference Contributions} \label{chapter:conferences}
\addcontentsline{toc}{chapter}{\nameref{chapter:conferences}}

\begin{itemize}

\item \textbf{Talk}: ``First-principles analysis of the efficiency of photovoltaic layers for CuAu-like chalcogenides and silicon'', at the \textit{International Conference on Advanced Materials Modelling}, September 5-7 2016 in Rennes, France.

\item \textbf{Poster}: ``First-Principles Investigation of Local Cation-Anion Interactions in Polyborane Solid Electrolytes'', at the \textit{Computational Chemistry and Materials Science Summer Institute}, June 5 - August 11 2017 in Lawrence Livermore National Laboratory, USA.

\item \textbf{Talk}: ``Exceeding the Shockley-Queisser limit within the detailed balance framework'', at the \textit{Deutsche Physikalische Gesellschaft Spring Meeting of the Condensed Matter Section}, March 11-16 2018 in Berlin, Germany.

\item \textbf{Talk}: ``First-principles modeling of secondary electron emission from slow ion bombardment'', at the \textit{Deutsche Physikalische Gesellschaft Spring Meeting of the Condensed Matter Section}, March 11-16 2018 in Berlin, Germany.

\item \textbf{Poster}: ``Exceeding the Shockley-Queisser limit within the detailed balance framework'', at the \textit{Belgian Physical Society meeting}, April 11 2018 in Antwerp, Belgium.

\item \textbf{Talk}: ``First-Principles Investigation of the Stability of the Oxygen Framework of Li-Rich Battery Cathodes", at the \textit{Materials Research Society fall meeting}, November 25-30 2018 in Boston, USA.

\item \textbf{Talk}: ``First-Principles Investigation of the Stability of the Oxygen Framework of Li-Rich Battery Cathodes", at the \textit{International Conference on Advanced Materials Modelling}, June 26-29 2019 in Rennes, France. 

\item \textbf{Poster}: ``First-Principles Investigation of the Stability of the Oxygen Framework of Li-Rich Battery Cathodes", at the \textit{Big Data Summer School}, September 9-13 2019 in Platja d'Aro, Spain.

\end{itemize}
