\chapter*{Publication list} \label{chapter:publications}
\addcontentsline{toc}{chapter}{\nameref{chapter:publications}}

\textbf{A1 - Journals}
\begin{itemize}
\item \textbf{M. Bercx}, N. Sarmadian, R. Saniz, B. Partoens and D. Lamoen, \href{https://www.doi.org/10.1039/C6CP03468C}{``First-principles analysis of the spectroscopic limited maximum efficiency of photovoltaic absorber layers for CuAu-like chalcogenides and silicon"}, \emph{Physical Chemistry Chemical  Physics} \textbf{18}, pp. 20542-20549 (2016).

\item T. Willhammar, K. Sentosun, S. Mourdikoudis, B. Goris, M. Kurttepeli, \textbf{M. Bercx}, D. Lamoen, B. Partoens, I. Pastoriza-Santos, J. P\'erez-Juste, L. M. Liz-Marz\'an, S. Bals, Gustaaf van Tendeloo, \href{https://doi.org/10.1038/ncomms14925}{``Structure and Vacancy Distribution in Copper Telluride Nanoparticles Determine Plasmon Activity in the Near Infrared"}, \emph{Nature Communications} \textbf{8}, 14925 (2017).

\item M. Dimitrievska, P. Shea, K. E. Kweon, \textbf{M. Bercx}, J. B. Varley, W. S. Tang, A. V. Skripov, V. Stavila, T. J. Udovic, and B. C. Wood, \href{https://doi.org/10.1002/aenm.201703422}{``Carbon Incorporation and Anion Dynamics as Synergistic Drivers for Ultrafast Diffusion in Superionic LiCB$_{11}$H$_{12}$ and NaCB$_{11}$H$_{12}$"}, \textit{Advanced Energy Materials} \textbf{8} (15), 1703422 (2018).

\item \textbf{M. Bercx}, B. Partoens, D. Lamoen, \href{https://doi.org/10.1103/PhysRevB.99.085413}{``Quantitative modeling of secondary electron emission from slow-ion bombardment on semiconductors"}, \textit{Physical Review B} \textbf{99}, 085413 (2019).

\item \textbf{M. Bercx}, L. Slap, B. Partoens and D. Lamoen, \href{https://doi.org/10.1557/adv.2019.135}{``First-Principles investigation of the oxygen framework of Li-rich battery cathodes"}, \textit{MRS Advances} \textbf{4}, pp. 813-820 (2019).

\item C. Callaert, \textbf{M. Bercx},  D. Lamoen and J. Hadermann, \href{https://doi.org/10.1107/S2052520619008357}{``Interstitial defects in the Van der Waals gap of Bi2Se3"}, \textit{Acta Crystallographica Section B} \textbf{75}, pp. 717-732 (2019).

\item K. Choudhary, \textbf{M. Bercx}, J. Jiang, R. Pachter, D. Lamoen and F. Tavazza, \href{https://doi.org/10.1021/ACS.CHEMMATER.9B02166
}{``Accelerated Discovery of Efficient Solar-cell Materials using Quantum and Machine-learning Methods"}, \textit{Chemistry of Materials} \textbf{31}, pp. 5900-5908 (2019). 

\item M. J{\o}rgensen, P. Shea, A. Tomich, J. Varley, \textbf{M. Bercx}, S. Lovera, R. \v{C}ern\'{y}, W. Zhou, T. J. Udovic, V. Lavallo, T. Jensen, B. Wood, V. Stavila, ``Understanding Superionic Conductivity in Lithium and Sodium Salts of Weakly Coordinating Closo-Hexahalocarbaborate Anions", \textit{under review with Chemistry of Materials}.

\end{itemize}

\textbf{B2 - Book chapter}

\begin{itemize}

\item \textbf{M. Bercx}, R. Saniz, B. Partoens, D. Lamoen, \href{https://link.springer.com/chapter/10.1007/978-3-319-72374-7_15}{``Exceeding the Shockley-Queisser limit within the detailed balance framework"}, In: Many-Body Approaches at Different Scales, ed. by G.G.N. Angilella and C. Amovilli; (New York : Springer, 2018), ISBN 978-3-319-72373-0 (2018).

\end{itemize}

\textbf{In preparation}

\begin{itemize}

\item A. Paulus, M. Hendrickx, \textbf{M. Bercx}, O. Karakulina, M. Kirsanova, D. Lamoen, A. M. Abakumov, J. Hadermann, M. K. Van Bael, and A. Hardy, ``An in-depth study of Sn substitution in Li-rich/Mn-rich NMC as a cathode material for Li-ion batteries", \textit{in preparation}.

\item \textbf{M. Bercx}, B. Partoens, D. Lamoen, ``Modeling secondary electron emission from slow-ion bombardment on metals by including plasmonic effects", \textit{in preparation}.

\end{itemize}
